---
title: "Task 5 - Right Wing Authoritarianism"
author: "Mandip"
date: "2023-04-16"
output: html_document
---

# Import libraries and data sets

### Import libraries

```{r}
library(assertr)
library(tidyverse)
```

### Import data with verification

- Check all fields required for analysis are present
- `person_id` not required

- Check `test_time`, `age`, `familysize` and `rwa` are numeric

```{r}
rwa_data <- read_csv(here::here("clean_data/rwa_clean.csv"),
                     show_col_types = FALSE)

rwa_data %>% 
  verify(has_all_names("test_time", "education", "childhood", "gender",
                       "age", "hand", "familysize", "rwa")) %>% 
  verify(is.numeric(c(test_time, age, familysize, rwa))) %>%
  summary()
```

```{r}
rwa_data %>% 
  head()

rwa_data %>% 
  summarise(across(.cols = everything(),
                   .fns = ~ sum(is.na(.x))))

rwa_data %>% 
  group_by(education) %>% 
  count()

rwa_data %>% 
  group_by(childhood) %>% 
  count()

rwa_data %>% 
  group_by(gender) %>% 
  count()

rwa_data %>% 
  group_by(hand) %>% 
  count()
```


# Analysis questions

### Q1
What’s the average RWA score for each gender?

```{r}
rwa_data %>% 
  filter(!is.na(gender)) %>% 
  summarise(average_RWA_score = mean(rwa, na.rm = TRUE),
            .by = gender)
```

### Q2
What’s the average RWA score for left handed people vs. right handed people?

```{r}
rwa_data %>%
  filter(!is.na(hand)) %>% 
  summarise(average_RWA_score = mean(rwa, na.rm = TRUE),
            .by = hand)
```

### Q3
What’s the average family size for each type of childhood?

```{r}
rwa_data %>% 
  filter(!is.na(childhood)) %>% 
  summarise(average_family_size = mean(familysize, na.rm = TRUE),
            .by = childhood)
```

### Q4
What’s the average time to take the test for each education level?

```{r}
rwa_data %>% 
  filter(!is.na(education)) %>% 
  summarise(average_time_in_seconds_to_take_test = mean(test_time, na.rm = TRUE),
            .by = education)
```

### Q5
Create a plot of results of question 4.

```{r}
rwa_data %>% 
  filter(!is.na(education)) %>% 
  summarise(average_time_in_seconds_to_take_test = mean(test_time, na.rm = TRUE),
            .by = education) %>% 
  ggplot(aes(x=factor(education,
                      levels = c("Less than high school",
                                 "High school",
                                 "University degree",
                                 "Graduate degree")),
             y=average_time_in_seconds_to_take_test,
             fill=education)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  geom_text(aes(label=round(average_time_in_seconds_to_take_test, 2)),
            vjust=2, color="black", show.legend = FALSE) +
  ggtitle("Average time to take test \n by Education level") +
  xlab("Education level") +
  ylab("Time (seconds)") +
  ylim(0, 250) +
  theme(plot.title = element_text(hjust=0.5, color="black", 
                                  size=14, face="bold.italic"),
        panel.grid.minor=element_blank())
```



### Q4
What’s the average RWA score for people aged:

- Under 18
- 18 to 25
- 26 to 40
- 41 to 60
- Over 60

```{r}
rwa_data %>% 
  filter(!is.na(age)) %>% 
  mutate(age_bracket = case_when(age < 18 ~ "Under 18",
                                 age <= 25 ~ "18 to 25",
                                 age <= 40 ~ "26 to 40",
                                 age <= 60 ~ "41 to 60",
                                 TRUE ~ "Over 60")) %>% 
  summarise(average_RWA_score = mean(rwa, na.rm = TRUE),
            .by = age_bracket)
```
