---
title: "Task 3 - Seabird observation data"
author: "Mandip"
date: "2023-04-14"
output: html_document
---

# Import libraries and data sets

### Import libraries

```{r}
library(tidyverse)
library(assertr)
```

### Import data with verification

- Check all fields required for analysis are present
  - `record`, `record_id` and `longitude` not required
  
- Check `count` and `latitude` are numeric

```{r}
seabirds_data <- read_csv(here::here("clean_data/seabirds_clean.csv"),
                           show_col_types = FALSE)
seabirds_data %>% 
  verify(has_all_names("common_name", "scientific_name", "abbreviation",
                       "count", "latitude")) %>% 
  verify(is.numeric(c(count, latitude))) %>%
  summary()
```

```{r}
seabirds_data %>% 
  group_by(common_name) %>% 
  count()

seabirds_data %>% 
  group_by(scientific_name) %>% 
  count()

seabirds_data %>% 
  group_by(abbreviation) %>% 
  count()
```

# Analysis questions

### Q1
Which bird had the most individual sightings?

```{r}
seabirds_data %>% 
  select(common_name, scientific_name, abbreviation) %>% 
  mutate(number_of_sightings = n(),
         .by = common_name) %>%
  unique() %>% 
  slice_max(number_of_sightings)
```

### Q2
Which bird had the highest total count?

```{r}
seabirds_data %>% 
  mutate(total_count = sum(count, na.rm = TRUE),
         .by = common_name) %>%
  select(common_name, scientific_name, abbreviation, total_count) %>%
  unique() %>% 
  slice_max(total_count)
```

### Q3
Which bird had the highest total count above a latitude of -30?

```{r}
seabirds_data %>% 
  filter(latitude > -30) %>% 
  mutate(total_count = sum(count, na.rm = TRUE),
         .by = common_name) %>%
  select(common_name, scientific_name, abbreviation, total_count) %>%
  unique() %>% 
  slice_max(total_count)
```

### Q4
How many different types of birds were only ever seen in groups of 1?

```{r}
seabirds_data %>% 
  group_by(common_name, scientific_name, abbreviation) %>% 
  summarise(max_group_number = max(count),
            .groups = "drop") %>%
  filter(max_group_number == 1)
```

### Q5
How many penguins were seen? (Hint: there are many types of penguin)

```{r}
seabirds_data %>% 
  filter(str_detect(common_name, "(?i)penguin")) %>% 
  summarise(number_penguin_sightings = n(),
            number_of_penguins_seen = sum(count, na.rm = TRUE))
```
