---
title: "Task 6 - Dog owners survey"
author: "Mandip"
date: "2023-04-13"
output: html_document
---

# Import libraries and data sets

### Import libraries

```{r}
library(assertr)
library(tidyverse)
```

### Import data with verification

- Check all fields required for analysis are present
  - `id`, `title` and `first_name` not required

- Check `amount_spent_on_dog_food` and `dog_age` are numeric

```{r}
survey_data <- read_csv(here::here("clean_data/dog_owners_survey_clean.csv"),
                           show_col_types = FALSE)

survey_data %>% 
  verify(has_all_names("last_name", "email", "amount_spent_on_dog_food",
                       "dog_size", "dog_gender", "dog_age")) %>% 
  verify(is.numeric(c(amount_spent_on_dog_food, dog_age))) %>%
  summary()
```

```{r}
survey_data %>% 
  head()

survey_data %>% 
  summarise(across(.cols = everything(),
                   .fns = ~ sum(is.na(.x))))

survey_data %>% 
  group_by(dog_size) %>% 
  count()

survey_data %>% 
  group_by(dog_gender) %>% 
  count()
```


# Analysis questions

### Q1
The client only counts a valid email address as one ending in ‘.com’. How many survey results have a valid email address?

```{r}
survey_data %>% 
  filter(str_detect(email, ".com$")) %>% 
  summarise(number_of_valid_email_addresses = n())
```

### Q2
What’s the average amount spent on dog food for each dog size?

```{r}
survey_data %>% 
  summarise(average_amount_spent = mean(amount_spent_on_dog_food, na.rm = TRUE),
            .by = dog_size) %>% 
  arrange(desc(average_amount_spent))
```

### Q3
For owners whose surname starts with a letter in the second half of the alphabet (N onwards) what is the average age of their dog?

```{r}
survey_data %>% 
  filter(str_detect(last_name, "^(?i)[N-Z]")) %>% 
  summarise(average_age_of_dog = mean(dog_age, na.rm = TRUE))
```

### Q4
The dog_age column is the age in dog years. If the conversion is 1 human year = 6 dog years, then what is the average human age for dogs of each gender?

```{r}
survey_data %>%
  mutate(human_age = dog_age / 6) %>% 
  summarise(average_human_age_of_dog = mean(human_age, na.rm = TRUE),
            .by = dog_gender)
```

### Q5
Create a plot of results of question 4.

```{r}
# using ggplot after filtering data
survey_data %>%
  mutate(human_age = dog_age / 6,
         dog_gender = case_when(is.na(dog_gender) ~ "Unknown",
                                TRUE ~ dog_gender)) %>% 
  summarise(average_human_age_of_dog = mean(human_age, na.rm = TRUE),
            .by = dog_gender) %>% 
  ggplot(aes(x=dog_gender, y=average_human_age_of_dog, fill=dog_gender)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  geom_text(aes(label=round(average_human_age_of_dog, 2)),
            vjust=2, color="black", show.legend = FALSE) +
  ggtitle("Plot of Average human age of dog \n by Gender") +
  xlab("Gender") +
  ylab("Average human age (years)") +
  ylim(0, 10) +
  theme(plot.title = element_text(hjust=0.5, color="black", 
                                  size=14, face="bold.italic"),
        panel.grid.minor=element_blank())
```


