---
title: "Task 2 - Cake ingredients"
author: "Mandip"
date: "2023-04-11"
output: html_document
---

# Import libraries and data sets

### Import libraries

```{r}
library(tidyverse)
library(assertr)
```

### Import data with verification

- Check all fields required for analysis are present

- Check `quantity` is numeric

```{r}
cake_data <- read_csv(here::here("clean_data/cake_ingredients_clean.csv"),
                           show_col_types = FALSE)

cake_data %>% 
  verify(has_all_names("cake", "ingredient", "quantity", "measure")) %>% 
  verify(is.numeric(quantity)) %>%
  summary()
```

```{r}
cake_data %>% 
  group_by(cake) %>% 
  count()

cake_data %>% 
  group_by(ingredient) %>% 
  count()

cake_data %>% 
  group_by(quantity) %>% 
  count()

cake_data %>% 
  group_by(measure) %>% 
  count()
```


# Analysis questions

### Q1
Which cake has the most cocoa in it?
```{r}
cake_data %>% 
  filter(ingredient == "Cocoa") %>%
  group_by(measure) %>% 
  slice_max(quantity)
```

### Q2
For sponge cake, how many cups of ingredients are used in total?

```{r}
cake_data %>% 
  filter(cake == "Sponge" & measure == "cup") %>% 
  summarise(quantity_of_cups_of_ingredients = sum(quantity, na.rm = TRUE))

# all units of measure
cake_data %>% 
  filter(cake == "Sponge") %>% 
  summarise(quantity_per_measure = sum(quantity, na.rm = TRUE),
            .by = measure)
```

### Q3
How many ingredients are measured in teaspoons?

```{r}
cake_data %>% 
  filter(measure == "teaspoon" & !is.na(quantity)) %>% 
  distinct(ingredient) %>% 
  arrange(ingredient)
```

### Q4
Which cake has the most unique ingredients?

```{r}
cake_data %>% 
  filter(!is.na(quantity)) %>%
  summarise(number_of_ingredients = n(),
            .by = cake) %>% 
  slice_max(number_of_ingredients, n = 1)
```

### Q5
Which ingredients are used only once?

```{r}
cake_data %>% 
  filter(!is.na(quantity)) %>% 
  count(ingredient) %>% 
  filter(n == 1)
```
