---
title: "Task 4 - Halloween Candy Data (All Data)"
author: "Mandip"
date: "2023-04-15"
output: html_document
---

# Import libraries and data sets

### Import libraries

```{r}
library(tidyverse)
library(assertr)

# removes: `summarise()` has grouped output by x
options(dplyr.summarise.inform = FALSE)
```

### Import data

- Check all fields required for analysis are present
- Check `age` is numeric
  - `year` can be character or numeric class

```{r}
candy_data <- read_csv(here::here("clean_data/candy_clean.csv"),
                           show_col_types = FALSE)

candy_data %>% 
  verify(has_all_names("year", "rater_id", "age", "gender", "country",
                       "going_trick_or_treating", "candy", "rating")) %>% 
  verify(is.numeric(age)) %>%
  summary()

candy_data_unique_rater <- candy_data %>% 
  select(year, rater_id, age, gender, country, going_trick_or_treating) %>% 
  unique()

candy_data_with_rating_numbers <- candy_data %>% 
  mutate(rating_number = case_when(rating == "despair" ~ -1,
                                   rating == "joy" ~ 1,
                                   TRUE ~ 0)
         )
```

```{r}
candy_data_unique_rater  %>% 
  group_by(year) %>% 
  count()

candy_data_unique_rater  %>%
  group_by(gender) %>% 
  count()

candy_data_unique_rater  %>%
  group_by(country) %>% 
  count()

candy_data_unique_rater  %>%
  group_by(going_trick_or_treating) %>% 
  count()

candy_data  %>% 
  group_by(candy) %>% 
  count()

candy_data  %>% 
  group_by(rating) %>% 
  count()
```

# Analysis questions

## Part 1

### Q1
What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

```{r}
# across all three years
candy_data  %>%
  summarise(number_of_ratings = n())

# split by year
candy_data  %>%
  summarise(number_of_ratings = n(),
            .by = year)
```

### Q2
What was the average age of people who are going out trick or treating?

```{r}
candy_data_unique_rater  %>%
  filter(going_trick_or_treating == "Yes") %>% 
  summarise(average_age = mean(age, na.rm = TRUE))
```

### Q3
What was the average age of people who are not going trick or treating?

```{r}
candy_data_unique_rater  %>%
  filter(going_trick_or_treating == "No") %>% 
  summarise(average_age = mean(age, na.rm = TRUE))
```

### Q4
For each of joy, despair and meh, which candy bar received the most of these ratings?

```{r}
# candy bars only
candy_data %>%
  filter(str_detect(candy, "bar")) %>% 
  group_by(rating, candy) %>% 
  summarise(number_of_ratings = n()) %>%
  slice_max(number_of_ratings, n = 1)

# all candy
candy_data %>%
  group_by(rating, candy) %>% 
  summarise(number_of_ratings = n()) %>%
  slice_max(number_of_ratings, n = 1)
```

### Q5
How many people rated Starburst as despair?

```{r}
candy_data %>% 
  filter(candy == "starburst" & rating == "despair") %>% 
  summarise(number_of_people_rated_starburst_as_despair = n())
```


## Part 2

For the next three questions, count despair as -1, joy as +1, and meh as 0

### Q6
What was the most popular candy bar by this rating system for each gender in the dataset ?

```{r}
# candy bars only
candy_data_with_rating_numbers %>%
  filter(str_detect(candy, "bar")) %>% 
  group_by(gender, candy) %>%
  summarise(rating_value = sum(rating_number, na.rm = TRUE)) %>% 
  slice_max(rating_value)

# all candy
candy_data_with_rating_numbers %>%
  group_by(gender, candy) %>%
  summarise(rating_value = sum(rating_number, na.rm = TRUE)) %>% 
  slice_max(rating_value)
```

### Q7
What was the most popular candy bar in each year?

```{r}
# candy bars only
candy_data_with_rating_numbers %>%
  filter(str_detect(candy, "bar")) %>% 
  group_by(year, candy) %>%
  summarise(rating_value = sum(rating_number, na.rm = TRUE)) %>% 
  slice_max(rating_value)

# all candy
candy_data_with_rating_numbers %>%
  group_by(year, candy) %>%
  summarise(rating_value = sum(rating_number, na.rm = TRUE)) %>% 
  slice_max(rating_value)
```


### Q8
What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

```{r}
# candy bars only
candy_data_with_rating_numbers %>%
  filter(str_detect(candy, "bar")) %>% 
  group_by(country, candy) %>%
  summarise(rating_value = sum(rating_number, na.rm = TRUE)) %>% 
  slice_max(rating_value)

# all candy
candy_data_with_rating_numbers %>%
  group_by(country, candy) %>%
  summarise(rating_value = sum(rating_number, na.rm = TRUE)) %>% 
  slice_max(rating_value)
```
