---
title: "Task 1 - Decathlon Data"
author: "Mandip"
date: "2023-04-11"
output: html_document
---

# Import libraries and data sets

### Import libraries

```{r}
library(tidyverse)
library(assertr)
```

### Import data with verification

- Check all fields required for analysis are present
  - `rank` not required
- Check `points` and `score` are numeric

```{r}
decathlon_data <- read_csv(here::here("clean_data/decathlon_clean.csv"),
                           show_col_types = FALSE)

decathlon_data %>% 
  verify(has_all_names("competitor", "points", "competition",
                       "event", "score")) %>% 
  verify(is.numeric(c(points, score))) %>%
  summary()
```

```{r}
decathlon_data %>% 
  group_by(competition) %>% 
  count()

decathlon_data %>% 
  group_by(event) %>% 
  count()
```

# Analysis questions

### Q1
Who had the longest long jump seen in the data?

```{r}
decathlon_data %>% 
  filter(event == "long jump") %>% 
  slice_max(score)
```

### Q2
What was the average 100m time in each competition?

```{r}
decathlon_data %>% 
  filter(event == "100m") %>% 
  summarise("average 100m time (s)" = mean(score),
            .by =competition)
```

### Q3
Who had the highest total points across both competitions?

```{r}
decathlon_data %>% 
  summarise(total_points = sum(points),
            .by = competitor) %>% 
  slice_max(total_points)
```

### Q4
What was the shot-put scores for the top three competitors in each competition?

```{r}
decathlon_data %>% 
  filter(event == "shot put") %>% 
  group_by(competition) %>%
  slice_max(score, n = 3)
```

### Q5
What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?

```{r}
# across both competitions
decathlon_data %>% 
  filter(event == "400m") %>% 
  mutate(less_than_50s = score < 50) %>%
  summarise(average_points = mean(points),
            .by = less_than_50s)

# for each competition
decathlon_data %>% 
  filter(event == "400m") %>% 
  mutate(less_than_50s = score < 50) %>%
  summarise(average_points = mean(points),
            .by = c(competition, less_than_50s))
```
